---
title: "Project 5 Digits"
author: "Po Yi Liu"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Library
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(vip)
library(parallel)
library(doParallel)
library(embed)
library(textrecipes)
library(xgboost)
```

## import data and target
```{r, message=FALSE, warning=FALSE}
digital_kaggle <- read_csv("D:/fallclass/Intro to Machine Learning/project5/digit_holdout.csv") %>%
  clean_names() 
digital_train <- read_csv("D:/fallclass/Intro to Machine Learning/project5/digit_train.csv") %>%
  clean_names()%>%
  mutate(label = factor(label))
head(digital_train)
```

## Train Test Split 
```{r, message=FALSE, warning=FALSE}
train_test_spit<- initial_split(digital_train, prop = 0.7)

train <- training(train_test_spit)
test  <- testing(train_test_spit)


# -- grey scale pixels are 0 - 255 
train_scaled <- train %>%
  mutate_if(is.numeric, funs(./255)) 

test_scaled <- test %>%
  mutate_if(is.numeric, funs(./255)) 

sprintf("Train PCT : %d", nrow(train))
sprintf("Test  PCT : %d", nrow(test))
train_cv_folds <- vfold_cv(train, v=5)
```
#define recipe for mlp
```{r, message=FALSE, warning=FALSE}
digital_recipe <- recipe(label ~ ., train_scaled)%>%
  step_rm(id)

```

#set mlp model
```{r, message=FALSE, warning=FALSE}
digital_mlp <- mlp(hidden_units = tune(), 
                   penalty = tune(), 
                   epochs = tune()) %>%
  set_engine("nnet", MaxNWts=10245) %>%
  set_mode("classification") 

digital_wf <- workflow() %>%
  add_recipe(digital_recipe) %>%
  add_model(digital_mlp)


```

#tunning mlp
```{r, message=FALSE, warning=FALSE}
# -- setup your tuning grid -- random force 
tune_grid_mlp <- grid_random(hidden_units(c(0,30)),
                    penalty(c(0.0, 1)),
                    epochs(c(10, 100)),
                    size = 15)
print(tune_grid_mlp)

# -- setup parallel process 
all_cores <- detectCores(logical = TRUE)
sprintf("# of Logical Cores: %d", all_cores)
cl <- makeCluster(all_cores)
registerDoParallel(cl)

# -- train!! K times for each parameter -- 
mlp_tuning_results <- digital_wf %>% 
  tune_grid(
    resamples = train_cv_folds,
    grid = tune_grid_mlp,
    control = control_resamples(save_pred = TRUE)
    )

mlp_tuning_results

```
#Review Tuning Results 
````{r, message=FALSE, warning=FALSE}
## -- results of tuning -- 
mlp_tuning_results %>% 
  collect_metrics() %>%
  mutate_if(is.numeric, round,3) %>% 
  pivot_wider(names_from = .metric, values_from=c(mean, std_err))
```
 
#selecting "best" parameters
```{r, message=FALSE, warning=FALSE}
mlp_tuning_results %>%
  show_best("accuracy") %>%
  print()

mlp_best <- mlp_tuning_results %>%
  select_best("accuracy") 

print(mlp_best)
```

#refitting workflow with "best" parameters
```{r, message=FALSE, warning=FALSE}
mlp_final_wf <- digital_wf %>% 
  finalize_workflow(mlp_best)

print(mlp_final_wf)

mlp_final_fit  <- mlp_final_wf %>%
  fit(data = train_scaled) 
```

#evaluate mlp performance
```{r, message=FALSE, warning=FALSE}
predict(mlp_final_fit, train_scaled,type="class") %>%
    bind_cols(., train )-> scored_train
    
predict(mlp_final_fit, test_scaled,type="class") %>%
    bind_cols(., test )-> scored_test

scored_test$label <- factor(scored_test$label,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

scored_test$.pred_class <- factor(scored_test$.pred_class,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

scored_train %>%
  mutate(.part = "train") %>%
  bind_rows(scored_test %>%
              mutate(.part = "test")
  ) %>%
  group_by(.part) %>%
    metrics(label, estimate = .pred_class) %>%
  filter(.metric == "accuracy") %>%
  spread(.metric, .estimate) %>%
  select(-.estimator)

 scored_test %>%
    conf_mat(label, .pred_class) %>%
    autoplot(type = "heatmap")
```

#prediction example when label=1
```{r, message=FALSE, warning=FALSE}
theme_set(theme_light())
theme_set(theme_light())

predict(mlp_final_fit, test_scaled,type="class") %>%
    bind_cols(., test )-> scored_test

x_gathered <- scored_test %>%
  filter(label == 1) %>%
  filter(label != .pred_class) %>%
  select(starts_with("x"), label,.pred_class ) %>%
  head(12) %>%
  mutate(instance = row_number()) %>%
  gather(x, value, -label, -instance, -.pred_class) %>%
  tidyr::extract(x, "x", "(\\d+)", convert = TRUE) %>%
  mutate(pixel = x - 2,
         x = pixel %% 28,
         y = 28 - pixel %/% 28)


x_gathered$label <- factor(x_gathered$label,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered$.pred_class <- factor(x_gathered$.pred_class,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))


x_gathered %>%
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~label + .pred_class)

```
#prediction example when label not equal to predict
```{r, message=FALSE, warning=FALSE}

predict(mlp_final_fit, test_scaled,type="class") %>%
    bind_cols(., test )-> scored_test

x_gathered <- scored_test %>%
  #filter(label == 0) %>%
  filter(label != .pred_class) %>%
  select(starts_with("x"), label,.pred_class ) %>%
  head(30) %>%
  mutate(instance = row_number()) %>%
  gather(x, value, -label, -instance, -.pred_class) %>%
  tidyr::extract(x, "x", "(\\d+)", convert = TRUE) %>%
  mutate(pixel = x - 2,
         x = pixel %% 28,
         y = 28 - pixel %/% 28)

x_gathered$label <- factor(x_gathered$label,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered$.pred_class <- factor(x_gathered$.pred_class,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered %>%
  ggplot(aes(x, y, fill = value)) +
  geom_raster() +
   scale_fill_gradient(low = "white", high = "black", na.value = NA) + 
  facet_wrap(~ label + .pred_class) 


```
#prediction example when label=0
```{r}
predict(mlp_final_fit, test_scaled,type="class") %>%
    bind_cols(., test )-> scored_test

x_gathered <- scored_test %>%
  filter(label == 0) %>% # cycle through examples 
  filter(label != .pred_class) %>%
  select(starts_with("x"), label,.pred_class ) %>%
  head(50) %>%
  mutate(instance = row_number()) %>%
  gather(x, value, -label, -instance, -.pred_class) %>%
  tidyr::extract(x, "x", "(\\d+)", convert = TRUE) %>%
  mutate(pixel = x - 2,
         x = pixel %% 28,
         y = 28 - pixel %/% 28)

x_gathered$label <- factor(x_gathered$label,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered$.pred_class <- factor(x_gathered$.pred_class,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered %>%
  ggplot(aes(x, y, fill = value)) +
  geom_raster() +
   scale_fill_gradient(low = "white", high = "black", na.value = NA) + 
  facet_wrap(~ label + .pred_class) 

```
#prediction example when label=0 to 9
```{r}
image_check <- function(label_id){
  predict(mlp_final_fit, test_scaled,type="class") %>%
    bind_cols(., test )-> scored_test

x_gathered <- scored_test %>%
  filter(label == label_id) %>% # cycle through examples 
  filter(label != .pred_class) %>%
  select(starts_with("x"), label,.pred_class ) %>%
  head(50) %>%
  mutate(instance = row_number()) %>%
  gather(x, value, -label, -instance, -.pred_class) %>%
  tidyr::extract(x, "x", "(\\d+)", convert = TRUE) %>%
  mutate(pixel = x - 2,
         x = pixel %% 28,
         y = 28 - pixel %/% 28)

x_gathered$label <- factor(x_gathered$label,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered$.pred_class <- factor(x_gathered$.pred_class,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered %>%
  ggplot(aes(x, y, fill = value)) +
  geom_raster() +
   scale_fill_gradient(low = "white", high = "black", na.value = NA) + 
  facet_wrap(~ label + .pred_class) 
  
}



for (i in range(0:9)){
  p <- image_check(i)
  print(p)
}

```

#define recipe for rf and xgboost
```{r, message=FALSE, warning=FALSE}
digital_recipe01 <- recipe(label ~ ., train)%>%
  step_rm(id)
```

```{r, message=FALSE, warning=FALSE}
bake_train <- bake(digital_recipe01%>%prep(), new_data = train)
```


## random forest
#Define the Model Document and hyper parameters
#Create a workflow and Fit the model
```{r, message=FALSE, warning=FALSE}
fraud_rf_spec <- rand_forest(
    trees  = tune(),
    min_n = tune(),
   ) %>% 
      set_engine("ranger", importance = "impurity") %>% 
      set_mode("classification")

fraud_rf_wf <- workflow() %>%
  add_recipe(digital_recipe01) %>%
  add_model(fraud_rf_spec) 
 

```


#tunning random forest
```{r, message=FALSE, warning=FALSE}
# -- setup your tuning grid -- random force 
tune_grid_rf <- grid_random(trees(c(100,500)),
                         min_n(),
                          size = 15)
print(tune_grid_rf)

# -- setup parallel process 
all_cores <- detectCores(logical = TRUE)
sprintf("# of Logical Cores: %d", all_cores)
cl <- makeCluster(all_cores)
registerDoParallel(cl)

# -- train!! K times for each parameter -- 
rf_tuning_results <- fraud_rf_wf %>% 
  tune_grid(
    resamples = train_cv_folds,
    grid = tune_grid_rf,
    control = control_resamples(save_pred = TRUE)
    )

rf_tuning_results

```

#Review Tuning Results 
````{r, message=FALSE, warning=FALSE}
## -- results of tuning -- 
rf_tuning_results %>% 
  collect_metrics() %>%
  mutate_if(is.numeric, round,3) %>% 
  pivot_wider(names_from = .metric, values_from=c(mean, std_err))
```

#Visualize impact 
```{r, message=FALSE, warning=FALSE}
## - visualize 
rf_tuning_results %>%
  collect_metrics() %>%
  mutate_if(is.numeric, round,3) %>%
  ggplot(aes(trees, mean, )) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 

rf_tuning_results %>%
  collect_metrics()  %>%
  mutate_if(is.numeric, round,3) %>%
  ggplot(aes(min_n, mean, )) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 

```


#selecting "best" parameters
```{r, message=FALSE, warning=FALSE}
rf_tuning_results %>%
  show_best("accuracy") %>%
  print()

rf_best <- rf_tuning_results %>%
  select_best("accuracy") 

print(rf_best)
```


#refitting workflow with "best" parameters
```{r, message=FALSE, warning=FALSE}
rf_final_wf <- fraud_rf_wf %>% 
  finalize_workflow(rf_best)

print(rf_final_wf)

rf_final_fit  <- rf_final_wf %>%
  fit(data = train) 
```

#model performance
```{r, message=FALSE, warning=FALSE}
# -- score training  
predict(rf_final_fit, train) %>%
  bind_cols(.,train)-> scored_train_rf 

# -- score testing 
predict(rf_final_fit, test) %>%
     bind_cols(., test) -> scored_test_rf   

# -- Metrics: Train and Test 
scored_train_rf %>% 
  metrics(label, .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_rf %>% 
               metrics(label, .pred_class) %>%
               mutate(part="testing") ) %>%
  pivot_wider(names_from = .metric, values_from=.estimate)
  
```


#prediction example when label not equal to predict
```{r, message=FALSE, warning=FALSE}
x_gathered_rf <- scored_test_rf %>%
  #filter(label == 0) %>%
  filter(label != .pred_class) %>%
  select(starts_with("x"), label,.pred_class ) %>%
  head(30) %>%
  mutate(instance = row_number()) %>%
  gather(x, value, -label, -instance, -.pred_class) %>%
  tidyr::extract(x, "x", "(\\d+)", convert = TRUE) %>%
  mutate(pixel = x - 2,
         x = pixel %% 28,
         y = 28 - pixel %/% 28)

x_gathered_rf$label <- factor(x_gathered_rf$label,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered_rf$.pred_class <- factor(x_gathered_rf$.pred_class,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered_rf %>%
  ggplot(aes(x, y, fill = value)) +
  geom_raster() +
   scale_fill_gradient(low = "white", high = "black", na.value = NA) + 
  facet_wrap(~ label + .pred_class) 


```

#prediction example when label= 0 to 9
```{r, message=FALSE, warning=FALSE}
image_check <- function(label_id){
  predict(digital_wf, test_scaled,type="class") %>%
    bind_cols(., test )-> scored_test_rf

x_gathered_rf01 <- scored_test_rf %>%
  filter(label == label_id) %>% # cycle through examples 
  filter(label != .pred_class) %>%
  select(starts_with("x"), label,.pred_class ) %>%
  head(50) %>%
  mutate(instance = row_number()) %>%
  gather(x, value, -label, -instance, -.pred_class) %>%
  tidyr::extract(x, "x", "(\\d+)", convert = TRUE) %>%
  mutate(pixel = x - 2,
         x = pixel %% 28,
         y = 28 - pixel %/% 28)

x_gathered_rf01$label <- factor(x_gathered_rf01$label,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered_rf01$.pred_class <- factor(x_gathered_rf01$.pred_class,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered_rf01 %>%
  ggplot(aes(x, y, fill = value)) +
  geom_raster() +
   scale_fill_gradient(low = "white", high = "black", na.value = NA) + 
  facet_wrap(~ label + .pred_class) 
  
}



for (i in range(0:9)){
  p <- image_check(i)
  print(p)
}

```

## xgboost
```{r, message=FALSE, warning=FALSE}
xgb_model <- boost_tree(
  trees = tune(), 
  min_n = tune(),            ## minimum number of observations 
  learn_rate = tune()        ## step size
) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")

xgb_model

# -- setup workflow 
xgb_workflow <- workflow() %>%
  add_recipe(digital_recipe01) %>%
  add_model(xgb_model) 

```

#tunning xgboost
```{r, message=FALSE, warning=FALSE}
tune_grid <- grid_random(trees(c(100,200)), 
                          min_n(),
                          learn_rate(),
                          size = 5)
print(tune_grid)
```

#tunning result
```{r, message=FALSE, warning=FALSE}
all_cores <- detectCores(logical = TRUE)
sprintf("# of Logical Cores: %d", all_cores)
cl <- makeCluster(all_cores)
registerDoParallel(cl)

xgb_tuning_results <- xgb_workflow %>%
  tune_grid(
  resamples = train_cv_folds,
  grid = tune_grid,
  control = control_resamples(save_pred = TRUE))
 
xgb_tuning_results
```


#Review Tuning Results 
````{r, message=FALSE, warning=FALSE}
## -- results of tuning -- 
xgb_tuning_results %>% 
  collect_metrics() %>%
  mutate_if(is.numeric, round,3) %>% 
  pivot_wider(names_from = .metric, values_from=c(mean, std_err))
```


#selecting "best" parameters
```{r, message=FALSE, warning=FALSE}
xgb_tuning_results %>%
  show_best("accuracy") %>%
  print()

xgb_best <- xgb_tuning_results %>%
  select_best("accuracy") 

print(xgb_best)
```

#refitting workflow with "best" parameters
```{r, message=FALSE, warning=FALSE}
xgb_final_wf <- xgb_workflow %>% 
  finalize_workflow(xgb_best)

print(xgb_final_wf)

xgb_final_fit  <- xgb_final_wf %>%
  fit(data = train) 
```

#model performance
```{r, message=FALSE, warning=FALSE}
# -- score training  
predict(xgb_final_fit, train) %>%
  bind_cols(.,train)-> scored_train_xgb 

# -- score testing 
predict(xgb_final_fit, test) %>%
     bind_cols(., test) -> scored_test_xgb   

# -- Metrics: Train and Test 
scored_train_xgb %>% 
  metrics(label, .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_xgb %>% 
               metrics(label, .pred_class) %>%
               mutate(part="testing") ) %>%
  pivot_wider(names_from = .metric, values_from=.estimate)
  
```

#prediction example when label not equal to predict
```{r, message=FALSE, warning=FALSE}
x_gathered_xgb <- scored_test_xgb %>%
  #filter(label == 0) %>%
  filter(label != .pred_class) %>%
  select(starts_with("x"), label,.pred_class ) %>%
  head(30) %>%
  mutate(instance = row_number()) %>%
  gather(x, value, -label, -instance, -.pred_class) %>%
  tidyr::extract(x, "x", "(\\d+)", convert = TRUE) %>%
  mutate(pixel = x - 2,
         x = pixel %% 28,
         y = 28 - pixel %/% 28)

x_gathered_xgb$label <- factor(x_gathered_xgb$label,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered_xgb$.pred_class <- factor(x_gathered_xgb$.pred_class,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered_xgb %>%
  ggplot(aes(x, y, fill = value)) +
  geom_raster() +
   scale_fill_gradient(low = "white", high = "black", na.value = NA) + 
  facet_wrap(~ label + .pred_class) 


```


#prediction example when label= 0 to 9
```{r, message=FALSE, warning=FALSE}
image_check <- function(label_id){
  predict(digital_wf, test_scaled,type="class") %>%
    bind_cols(., test )-> scored_test_xgb

x_gathered_xgb01 <- scored_test_xgb %>%
  filter(label == label_id) %>% # cycle through examples 
  filter(label != .pred_class) %>%
  select(starts_with("x"), label,.pred_class ) %>%
  head(50) %>%
  mutate(instance = row_number()) %>%
  gather(x, value, -label, -instance, -.pred_class) %>%
  tidyr::extract(x, "x", "(\\d+)", convert = TRUE) %>%
  mutate(pixel = x - 2,
         x = pixel %% 28,
         y = 28 - pixel %/% 28)

x_gathered_xgb01$label <- factor(x_gathered_xgb01$label,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered_xgb01$.pred_class <- factor(x_gathered_xgb01$.pred_class,
levels = c(0,1,2,3,4,5,6,7,8,9))
#labels = c("T-shirt/top", "Trouser", "Pullover", "Dress", "Coat", "Sandal", "Shirt", "Sneaker", "Bag", "Ankle boot"))

x_gathered_xgb01 %>%
  ggplot(aes(x, y, fill = value)) +
  geom_raster() +
   scale_fill_gradient(low = "white", high = "black", na.value = NA) + 
  facet_wrap(~ label + .pred_class) 
  
}



for (i in range(0:9)){
  p <- image_check(i)
  print(p)
}

```

